---
title: "Movement analysis"
author: "Jack B"
date: "2024-11-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Housekeeping #Note: not all packages are required
```{r, include=FALSE}
library(tidyverse)
library(sf)
library(move2)
library(units)
library(data.table)
```


Load data
```{r}
muddyfoot_sub <- readRDS("C:/Users/jacka/Documents/Academia/Projects/Landscape of fear - personality - Jack/Mortality/03_muddyfoot_sub.rds")
bt_sub <- readRDS("C:/Users/jacka/Documents/Academia/Projects/Landscape of fear - personality - Jack/Mortality/03_lake_BT_sub.rds")
cow_sub <- readRDS("C:/Users/jacka/Documents/Academia/Projects/Landscape of fear - personality - Jack/Mortality/03_lake_cow_sub.rds")
```

Combine data
```{r}
combined <- rbind(muddyfoot_sub, bt_sub, cow_sub)
#Remove old data files
rm(muddyfoot_sub, bt_sub, cow_sub)
#clear memory
gc()
```

Filter out pike
```{r}
combined <- combined %>% dplyr::filter(Species != 'Pike') 
```

Filter for individuals that were not found alive at the end of the expeirment
```{r}
suspect_mort = fread("C:/Users/jacka/Documents/Academia/Projects/Landscape of fear - personality - Jack/Mortality/suspected_mortality.csv", header = TRUE)
head(suspect_mort)
```

Filter for  suspected dead individuals
```{r}
mortality <- combined %>%  dplyr::filter(individual_ID %in% suspect_mort$individual_ID)
rm(combined)
```

Arrange by individual and datetime
```{r}
mortality <- mortality %>%
  arrange(individual_ID,timestamp)
```


Reduce influence of positional error

Using average positional estimates
```{r}
#Creating averaged positional estimates for each minute
average_positions <- mortality %>%
  mutate(time_1min = floor_date(timestamp, "1 minute")) %>% 
  group_by(individual_ID, time_1min) %>%
  mutate(avg_longitude = mean(longitude, na.rm = TRUE),
         avg_latitude = mean(latitude, na.rm = TRUE)) %>% 
  ungroup() %>%
  unite(indiv_time, c("individual_ID", "time_1min"), remove = FALSE) %>%
  distinct(indiv_time, .keep_all = T) %>% select(-c(longitude, latitude))
rm(mortality)
```

Convert to move object
```{r}
combined_mv <- mt_as_move2(average_positions, time_column = 'time_1min', track_id_column = 'individual_ID',
                                  coords = c("avg_longitude", "avg_latitude")) %>% sf::st_set_crs(4326) %>% st_transform(32634)
rm(average_positions)
```


Trajectory analysis
```{r}
# Calculate speed and distance
combined_mv <- combined_mv %>% mutate(speed = mt_speed(combined_mv),
                      distance = mt_distance(combined_mv))

#Change units
combined_mv$speed <- units::set_units(combined_mv$speed, cm/s)
combined_mv$distance <- units::set_units(combined_mv$distance, m)

#Convert to data frame
combined_mv <- as.data.frame(combined_mv)
```

Filter out unrealistic speeds (based on data from https://doi.org/10.1038/s41598-020-75974-x)
```{r}
#Load Ucrit data from https://doi.org/10.1038/s41598-020-75974-x
ucrit_data = fread("C:/Users/jacka/Documents/Academia/Projects/Landscape of fear - personality - Jack/data/ucrit_data.csv", header = TRUE)

ucrit <- ucrit_data %>% dplyr::filter(Species == 'Perca fluviatilis' | Species == 'Rutilus rutilus') %>% 
  group_by(Species) %>%
  dplyr::summarise(mean = mean(`Ucrit (cm/s)`))
ucrit
#Going to filter by the mean UCrit value = 97.72 cm/s for perch, and 84.075 cm/s for roach
```

Filter out unrealistic speeds
```{r}
combined_mv <- combined_mv %>% 
  dplyr::filter((Species == "Perch" & speed <= units::set_units(97.72, "cm/s")) |
                (Species == "Roach" & speed <= units::set_units(84.075, "cm/s")))
```


Summarise average daily speed
```{r}
#Summarise per day
daily_data <- combined_mv %>%
  dplyr::group_by(individual_ID, Date) %>%
  dplyr::mutate(
    avg_daily_speed_cm = mean(speed),
    avg_daily_speed_bl = as.numeric(mean(speed/Total_length)),
    total_distance = sum(distance)) %>% 
  dplyr::distinct(individual_ID, Date, .keep_all = TRUE) %>%
  ungroup()
```

What is the distribution of total daily distance? Bottom 5 % of total daily distance when fish were presumed alive is ~100 m per day
```{r}
daily_data %>% dplyr::filter(between(Date, '2022-09-26','2022-09-30')) %>% #first five days of tracking that is consistent between lakes
  dplyr::summarise(mean = mean(total_distance, na.rm = T),
                   sd = sd(total_distance, na.rm = T),
                   max = max(total_distance, na.rm = T),
                   min = min(total_distance, na.rm = T),
                   quartile_05 = quantile(total_distance, na.rm = T, probs = c(0.05))) %>%
  ungroup()
```

Making plots for each individual
```{r}
unique(daily_data$individual_ID)

#Make distance numeric 
daily_data$total_distance <- as.numeric(daily_data$total_distance)

#Create list of individuals
individual_id <- as.list(unique(daily_data$individual_ID))

# Create an empty list to store results
results_list <- list()

# Loop over each individual
for (i in seq_along(individual_id)) {
  tag <- individual_id[[i]]
  
  # Filter data for the current individual
  individual_data <- daily_data %>% filter(individual_ID == tag)
  
  # Run plot
  plot <- individual_data %>%
  ggplot(mapping = aes(x = Date, y = total_distance, color = individual_ID)) +
  geom_line(linewidth = 1) + 
  labs(y = "Total distance per day", x = "Date") +
  geom_point(aes(shape = as.factor(poor_tracking_day)), size = 2, color = 'black') +
  theme_bw() +
  scale_y_continuous(
    breaks = seq(0, max(individual_data$total_distance, na.rm = TRUE), by = 200),
    limits = c(0, max(individual_data$total_distance, na.rm = TRUE))
  ) +
  geom_hline(yintercept = 100, linetype = "dotted", color = "red")
  
  # Save results to the list with indexed label
  results_list[[paste0("plot_", i)]] <- plot
  
  # Save the plot as a PNG file
  ggsave(filename = paste0("C:/Users/jacka/Documents/Academia/Projects/Landscape of fear - personality - Jack/Mortality/Plots/", tag, ".png"), plot = plot)
  
  # Remove the object from the environment
  rm(individual_data, plot)
}
```

Which fish were known to be predated?
```{r}
daily_data %>%
  dplyr::distinct(individual_ID, .keep_all = TRUE) %>%
  dplyr::filter(Known_predated == 1)
```

